---
title: "p8105_hw1_yx2711"
author: "Yingchen Xu"
date: "2022-09-23"
output: github_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Problem 1
```{r}
data("penguins", package = "palmerpenguins")
penguins = penguins
# find number of columns and rows for penguins dataset
ncol(penguins)
nrow(penguins)
# find mean of flipper length
summary(penguins$flipper_length_mm)
```

There are 344 observations and 8 variables in the penguins dataset. The variables are species, island, bill length, bill depth, flipper length, body mass, sex and year. There are 8 columns and 344 rows. The mean of flipper length is 200.9.

```{r}
# Make a scatterplot for bill length vs flipper length, which is group by species.
scatterplot = ggplot(data = penguins, aes(x = bill_length_mm, y = flipper_length_mm)) + geom_point() + aes(color = species)

scatterplot

# Save scatterplot
ggsave("scatterplot.pdf")
```

# Problem 2
```{r}
# Create a dataframe called df with a random sample of size 10 from a standard Normal distribution, a logical vector, a character vector and a factor vector
df = 
  tibble(
    var_norm = rnorm(n = 10),
    var_log = c(var_norm > 0),
    var_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    var_fac = factor(c('low','med','high','low','med','high','low','med','high','low'))
  )

# Use the pull function to calculate the mean of variables in the data frame df.
mean(pull(df, 1))
mean(pull(df, 2))
mean(pull(df, 3))
mean(pull(df, 4))
```

The calculations of mean works for the random sample from the standard normal distribution and the logical vector. It does not work for the character vector and the factor vector.

```{r, eval = FALSE}
as.numeric(pull(df, 2))
as.numeric(pull(df, 3))
as.numeric(pull(df, 4))
```

After converting the variables into numeric type, the logical variable turns in to a group of 1 and 0.The character variable turns in to a group of NA, and the factor variable turns in to 1, 2, 3 depending on its level.
When using the pull function to calculate the mean, the warning suggests that the argument is not numeric or logical. The mean of logical variable could be calculated because logical variable exists as 1 or 0. The mean of character variable and factor variable cannot be calculated because they are not stored as any numeric format. 